---
title: "Biology Predictions"
author: "Jenny Rogers"
date: "April 30, 2020"
output: html_document
---

```{r}

library(tidyverse)
library(sf)
library(lubridate)
library(RColorBrewer)
library(gridExtra)
library(randomForest)


```

In this document, we make predictions for the biology using the stream temp and the hydrologic metrics and the already built biological models.

first, we build a model to predict spp presence or absence with the temperature data. we run PCA on the training data so that we run a logistic regression with spp presence and absece data using PCA predictor variables. This lets us use all the temp varialbes for each spp, but ensures that the variables are not correlated.  Then we run the logistic regression..
```{r}

#load("L:/Flow ecology and climate change_ES/Jenny/AirTemp/Modeling/spp_tmp_train_dat.RData")
#save(species, file = "spp_tmp_dat.RData")
load("spp_tmp_dat.RData")
pca <- prcomp(species[,5:10], scale=TRUE)
summary(pca)

write.csv(round(pca$rotation, 2), file = "pcaRotation.csv")

spp <- cbind(species, pca$x)
spp <- spp %>% 
  select(name, occurrence, PC1, PC2)

trt <- spp %>% filter(name %in% "rainbow trout")
suc <- spp %>% filter(name %in% "santa ana sucker")
vir <- spp %>% filter(name %in% "least bell's vireo")
chb <- spp %>% filter(name %in% "arroyo chub")
toa <- spp %>% filter(name %in% "arroyo toad")
tur <- spp %>% filter(name %in% "southwestern pond turtle")


summary(trout_mdl <- glm(occurrence ~ PC1+PC2, family = "binomial",data = trt))
summary(sucker_mdl <- glm(occurrence ~ PC1+PC2, family = "binomial",data = suc))
summary(vireo_mdl <- glm(occurrence ~ PC1+PC2, family = "binomial",data = vir))
summary(chub_mdl <- glm(occurrence ~ PC1+PC2, family = "binomial",data = chb))
summary(toad_mdl <- glm(occurrence ~ PC1+PC2, family = "binomial",data = toa))
summary(turtle_mdl <- glm(occurrence ~ PC1+PC2, family = "binomial",data = tur))



```

In this code chunk, we read in the random forest models for each spp to predict P(occurrence) based on streamflow, and the streamflow metric and temperature dataframes


```{r}
#models based on streamflow metrics
load("biology_model/mod_chub_rf.RData")
load("biology_model/mod_vireo_rf.RData")
load("biology_model/mod_toad_rf.RData")
load("biology_model/mod_turtle_rf.RData")
load("biology_model/mod_sucker_rf.RData")
load("biology_model/mod_trout_rf.RData")


#load clusters and elevation data
clusters<- st_read("C:/Users/JennyT/Documents/LitReview/RB4/StreamCat/COMID clustering.shp") %>% 
  data.frame() %>% 
  select(COMID, clstrCt, dam) %>% 
  filter(clstrCt == 1 & dam == 0 )

load("elv_comid.RData")

#streamflow metric data
#baseline flow
load("bsflowmetest.RData")
baseline_flow <- bsflowmetest %>% 
  select(var, COMID, dtsl, est) %>% 
  filter(COMID %in% clusters$COMID) %>% 
  mutate(year = year(dtsl)) %>% 
  spread(var, est)

#future flow
load("flowmetdt2.RData")
future_flow <- flowmetdt2 %>% 
  select(var, COMID, dtsl, est) %>% 
  filter(COMID %in% clusters$COMID) %>% 
  mutate(year = year(dtsl)) %>% 
  spread(var, est)

flow <- bind_rows(baseline_flow, future_flow)
flow <- left_join(flow, elv, by = "COMID")

#temperature data
load("fut_strm_tmp_wtrshd.RData")
load("baseline_stream_temp_wtrshd.RData")
temp <- rbind(baseline_stream_temp, future) %>% 
  filter(COMID %in% clusters$COMID) #filter to remove COMIDs in urban areas

#filter to make the COMIDs line up with the temp (likely removing the coastline)
flow <- flow %>% filter(COMID %in% temp$COMID) 

rm(baseline_stream_temp, future, bsflowmetest, flowmetdt2, clusters, baseline_flow, future_flow, elv)

```


In this code chunk, we transform the temperature data for prediction into the PC1 and PC2 variables
```{r}


pred_tmp <- temp[,3:8]

std_dat <- scale(pred_tmp)

PC1 <- as.matrix(std_dat) %*% as.vector(pca$rotation[,1])

PC2 <- as.matrix(std_dat) %*% as.vector(pca$rotation[,2])

rm(pred_tmp)

temp <- cbind(temp, PC1, PC2)

```

Here we will use the logistic reg models for spp presence and absense based on the temperature PCA models to make predictions for Temp and add in a column for baseline, mid-century, and end-of-centry
```{r}

temp$trout <- predict(trout_mdl, newdata = temp, type = "response")
temp$chub <- predict(chub_mdl, newdata = temp, type = "response")
temp$sucker <- predict(sucker_mdl, newdata = temp, type = "response")
temp$toad <- predict(toad_mdl, newdata = temp, type = "response")
temp$vireo <- predict(vireo_mdl, newdata = temp, type = "response")
temp$turtle <- predict(turtle_mdl, newdata = temp, type = "response")

temp$time <- ifelse(temp$year<=2014, "baseline", 
                    ifelse(temp$year >= 2022 & temp$year <= 2061, "mid", "end"))

```

Because PCA is hard to interpret, we will just plot each temp metric with prob of occurrence so people can see relatipmship
```{r}
t <- temp  %>% 
  rename("Max 7-Day Mean" = "Max_Wkly_Mean_StreamT",
         "Max 7-Day Max" = "Max_Wkl_Max_StreamT",
         "Min 7-Day Min" = "Min_Wkl_Min_StreamT",
         "Max 7-Day Rng" = "Max_Wkl_Rng_StreamT",
         "Mean 7-Day Rng" = "Mean_Wkl_Rng_StreamT") %>% 
  gather(3:8, key = "metric", value = "Temp") %>% 
  gather(8:13, key = "species", value = "Probablilty") %>% 
  filter(metric != "Max_Wkl_Max_StreamT_grt_30_")

p <- ggplot(data = t, mapping = aes(x = Temp, y = Probablilty, color = metric))+
  geom_smooth(method = "glm", method.args = list(family = "binomial")) +
  theme(panel.grid.major=element_line(colour="grey90"),
        panel.background = element_blank(),
        axis.title.x = element_text(size=30), 
        axis.title.y = element_text(size = 30),
        axis.text.x = element_text(hjust = 1, size = 20),
        axis.text.y = element_text(size = 30),
        title = element_text(size = 30),
        strip.text = element_text(size = 30, face = "bold"),
        strip.background = element_rect(fill = "lightblue", colour = "black"),
        legend.title = element_blank(),
        legend.text=element_text(size=30),
        legend.key.size = unit(3,"line"))+
  labs(x = "Temperature (°C)", y = "Probability")+
  facet_wrap(vars(species), labeller = as_labeller(c("chub" = "Arroyo chub", "sucker" = "Santa Ana sucker",
                                                               "toad" = "Arroyo toad", "trout" = "Rainbow trout",
                                                               "turtle" = "SW pond turtle", "vireo" = "Least Bell's vireo"))) 

ggsave("spp_prob_response_to_Temp.tiff", plot = p, dpi = 300, width = 20, height = 10, compression = "lzw")


```

In this chunk we will make a  figure of line graph showing mean and 5th 9th percentile probablilities in all years faceting for each spp.

```{r}

#plot for average (+high and low percentiles) annual probabilities across all COMID
timeline <- temp %>% 
  select(COMID, year, elv, trout, chub, sucker, toad, vireo, turtle) %>% 
  pivot_longer(cols = 4:9, names_to = "Species", values_to = "Probability") %>% 
  group_by(year, Species) %>% 
  summarise(avg = mean(Probability),
            upr_95 = quantile(Probability, probs = .95),
            lwr_05 = quantile(Probability, probs = .05)
    
  )


# expand yr with full NA values ---- Not Working!!!
# timeline <- seq(min(timeline$year), max(timeline$year)) %>%
#   tibble(
#     year = .
#   ) %>%
#   left_join(timeline, by = 'year')



p <- ggplot(data = timeline, mapping = aes(x = year, y = avg)) + 
  geom_line(lwd = 2)+
  geom_smooth(method = "lm")+
  geom_line(data = timeline, mapping = aes(x = year, y = upr_95), color = "grey60")+
  geom_line(data = timeline, mapping = aes(x = year, y = lwr_05), color = "grey60")+
  theme(panel.grid.major=element_line(colour="grey90"),
        panel.background = element_blank(),
        axis.title.x = element_text(size=30), 
        axis.title.y = element_text(size = 30),
        axis.text.x = element_text(hjust = 1, size = 20),
        axis.text.y = element_text(size = 20),
        title = element_text(size = 30),
        strip.text = element_text(size = 30, face = "bold"),
        strip.background = element_rect(fill = "lightblue", colour = "black"))+
  scale_x_continuous(breaks = seq(1980, 2100, by = 20))+
  scale_y_continuous(breaks = seq(0, 1, by = 0.25))+
  labs(x = NULL, y = "Probability")+
  facet_wrap(vars(Species), nrow = 3, labeller = as_labeller(c("chub" = "Arroyo chub", "sucker" = "Santa Ana sucker",
                                                               "toad" = "Arroyo toad", "trout" = "Rainbow trout",
                                                               "turtle" = "SW pond turtle", "vireo" = "Least Bell's vireo")))
 
  
ggsave("spp_prob_timeseries.tiff", plot = p, dpi = 300, width = 25, height = 20, compression = "lzw")

#get slopes for each spp so we can comment on the rate of change
c <- summary(lm(avg ~ year, data = timeline[timeline$Species == "chub",]))
d <- summary(lm(avg ~ year, data = timeline[timeline$Species == "trout",]))
e <- summary(lm(avg ~ year, data = timeline[timeline$Species == "vireo",]))
f <- summary(lm(avg ~ year, data = timeline[timeline$Species == "toad",]))
g <- summary(lm(avg ~ year, data = timeline[timeline$Species == "turtle",]))
h <- summary(lm(avg ~ year, data = timeline[timeline$Species == "sucker",]))

t(data.frame(
  "chub slope" = coefficients(c)[2,1],
  "trt slope" = coefficients(d)[2,1],
  "vir slope" = coefficients(e)[2,1],
  "toa slope" = coefficients(f)[2,1],
  "tur slope" = coefficients(g)[2,1],
  "suc slope" = coefficients(h)[2,1]
  
))

```


violine plots in change of the dist of change in prob from baseline to future for high and low elv  for each spp and a map showing change for each spp.
```{r}


df<- temp %>%
  select(COMID, year, elv, trout, chub, sucker, toad, vireo, turtle, time) %>% 
  filter(time %in% c("baseline", "end")) %>% 
  pivot_longer(cols = 4:9, names_to = "Species", values_to = "Probability") %>%
  group_by(Species, time, COMID)%>% 
  summarise(avgProb = mean(Probability)) %>% 
  ungroup()

chng <- 
  data.frame(
    "COMID" = unique(df$COMID),
    "chub_chg" = df$avgProb[df$Species == "chub" & df$time == "end"] - df$avgProb[df$Species == "chub" & df$time == "baseline"],
    "suc_chg" = df$avgProb[df$Species == "sucker" & df$time == "end"] - df$avgProb[df$Species == "sucker" & df$time == "baseline"],
    "trt_chg" = df$avgProb[df$Species == "trout" & df$time == "end"] - df$avgProb[df$Species == "trout" & df$time == "baseline"],
    "vir_chg" = df$avgProb[df$Species == "vireo" & df$time == "end"] - df$avgProb[df$Species == "vireo" & df$time == "baseline"],
    "toa_chg" = df$avgProb[df$Species == "toad" & df$time == "end"] - df$avgProb[df$Species == "toad" & df$time == "baseline"],
    "tur_chg" = df$avgProb[df$Species == "turtle" & df$time == "end"] - df$avgProb[df$Species == "turtle" & df$time == "baseline"]
    )
#calculate row averages and percentiles to get the average change across GCM and COMID in each metric
round(colMeans(chng),2)

round(apply(chng, 2, function (x){
  quantile(x, probs = 0.95)
}), 2)

round(apply(chng, 2, function (x){
  quantile(x, probs = 0.05)
}), 2)

#attach to spatial file
load("NHD_strm_rvr_map.RData")
chng <- left_join(NHD, chng, by = "COMID") %>% 
  filter(!is.na(chub_chg))

load("wtrshd_bundry.RData")
wtrshd_bndry<- st_transform(wtrshd_bndry, crs = st_crs(chng))

myPalette <- colorRampPalette(brewer.pal(9, "RdYlBu"))

a <- ggplot(data = chng, aes(col=chng$chub_chg)) +
  geom_sf(data = wtrshd_bndry, fill = "white", color = gray(.3), lwd = 1) +
  geom_sf()+
  scale_color_gradientn(colours = myPalette(100), limits = c(min(chng$chub_chg), max(chng$chub_chg)))+
  labs(col = "\u0394 P") +  ggtitle("(A)", subtitle = "Arroyo chub") + 
  theme(panel.grid.major=element_line(colour="transparent"),panel.background = element_blank(),
        axis.text = element_blank(),axis.ticks = element_blank(),
        text = element_text(size=30))
b <- ggplot(data = chng, aes(col=chng$suc_chg)) +
  geom_sf(data = wtrshd_bndry, fill = "white", color = gray(.3), lwd = 1) +
  geom_sf()+
  scale_color_gradientn(colours = myPalette(100), limits = c(min(chng$suc_chg), max(chng$suc_chg)))+
  labs(col = "\u0394 P") +  ggtitle("(B)", subtitle = "Santa Ana sucker") + 
  theme(panel.grid.major=element_line(colour="transparent"),panel.background = element_blank(),
        axis.text = element_blank(),axis.ticks = element_blank(),
        text = element_text(size=30))
c <- ggplot(data = chng, aes(col=chng$trt_chg)) +
  geom_sf(data = wtrshd_bndry, fill = "white", color = gray(.3), lwd = 1) +
  geom_sf()+
  scale_color_gradientn(colours = myPalette(100), limits = c(min(chng$trt_chg), max(chng$trt_chg)))+
  labs(col = "\u0394 P") +  ggtitle("(C)", subtitle = "rainbow trout") + 
  theme(panel.grid.major=element_line(colour="transparent"),panel.background = element_blank(),
        axis.text = element_blank(),axis.ticks = element_blank(),
        text = element_text(size=30))
d <- ggplot(data = chng, aes(col=chng$vir_chg)) +
  geom_sf(data = wtrshd_bndry, fill = "white", color = gray(.3), lwd = 1) +
  geom_sf()+
  scale_color_gradientn(colours = myPalette(100), limits = c(min(chng$vir_chg), max(chng$vir_chg)))+
  labs(col = "\u0394 P") +  ggtitle("(D)", subtitle = "Least Bell's vireo") + 
  theme(panel.grid.major=element_line(colour="transparent"),panel.background = element_blank(),
        axis.text = element_blank(),axis.ticks = element_blank(),
        text = element_text(size=30))
e <- ggplot(data = chng, aes(col=chng$toa_chg)) +
  geom_sf(data = wtrshd_bndry, fill = "white", color = gray(.3), lwd = 1) +
  geom_sf()+
  scale_color_gradientn(colours = myPalette(100), limits = c(min(chng$toa_chg), max(chng$toa_chg)))+
  labs(col = "\u0394 P") +  ggtitle("(E)", subtitle = "Arroyo toad") + 
  theme(panel.grid.major=element_line(colour="transparent"),panel.background = element_blank(),
        axis.text = element_blank(),axis.ticks = element_blank(),
        text = element_text(size=30))
f <- ggplot(data = chng, aes(col=chng$tur_chg)) +
  geom_sf(data = wtrshd_bndry, fill = "white", color = gray(.3), lwd = 1) +
  geom_sf()+
  scale_color_gradientn(colours = myPalette(100), limits = c(min(chng$tur_chg), max(chng$tur_chg)))+
  labs(col = "\u0394 P") +  ggtitle("(F)", subtitle = "Southwestern pond turtle") + 
  theme(panel.grid.major=element_line(colour="transparent"),panel.background = element_blank(),
        axis.text = element_blank(),axis.ticks = element_blank(),
        text = element_text(size=30))



comid_elv <- temp %>% 
  select(COMID, elv) %>% unique
chng <- left_join(chng, comid_elv, by = "COMID") %>% 
  data.frame() %>% 
  select(-geometry)

round(colMeans(chng[chng$elv=="high", 2:7]),2)
round(colMeans(chng[chng$elv=="low", 2:7]),2)
t.test(chng[chng$elv=="low", 2], chng[chng$elv=="high", 2])#chub
t.test(chng[chng$elv=="low", 3], chng[chng$elv=="high", 3])#suc
t.test(chng[chng$elv=="low", 4], chng[chng$elv=="high", 4])#trt
t.test(chng[chng$elv=="low", 5], chng[chng$elv=="high", 5])#vir
t.test(chng[chng$elv=="low", 6], chng[chng$elv=="high", 6])#toa
t.test(chng[chng$elv=="low", 7], chng[chng$elv=="high", 7])#tur

g <- ggplot(chng, aes(x = factor(elv, levels = c("low", "high")), y = chub_chg))+
  geom_violin()+geom_boxplot(width = .08)+
  labs(y="\u0394 P", x = "Elevation group")+
  ggtitle("(G)", subtitle = "Arroyo chub")+ 
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5))
h <- ggplot(chng, aes(x = factor(elv, levels = c("low", "high")), y = suc_chg))+
  geom_violin()+geom_boxplot(width = .08)+
  labs(y="\u0394 P", x = NULL)+
  ggtitle("(H)", subtitle = "Santa Ana sucker")+ 
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5))
i <- ggplot(chng, aes(x = factor(elv, levels = c("low", "high")), y = trt_chg))+
  geom_violin()+geom_boxplot(width = .08)+
  labs(y="\u0394 P", x = NULL)+
  ggtitle("(I)", subtitle = "rainbow trout")+ 
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5))
j <- ggplot(chng, aes(x = factor(elv, levels = c("low", "high")), y = vir_chg))+
  geom_violin()+geom_boxplot(width = .08)+
  labs(y="\u0394 P", x = NULL)+
  ggtitle("(J)", subtitle = "Least Bell's vireo")+ 
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5))
k <- ggplot(chng, aes(x = factor(elv, levels = c("low", "high")), y = toa_chg))+
  geom_violin()+geom_boxplot(width = .08)+
  labs(y="\u0394 P", x = NULL)+
  ggtitle("(K)", subtitle = "Arroyo toad")+ 
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5))
l <- ggplot(chng, aes(x = factor(elv, levels = c("low", "high")), y = tur_chg))+
  geom_violin()+geom_boxplot(width = .08)+
  labs(y="\u0394 P", x = NULL)+
  ggtitle("(L)", subtitle = "southwestern pond turtle")+ 
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5))

margin = theme(plot.margin = unit(c(5,5,5,5), "mm"))
p<-arrangeGrob(
  a + margin, g + margin, 
  b + margin, h + margin, 
  c + margin, i + margin, 
  d + margin, j + margin, 
  e + margin, k + margin, 
  f + margin, l + margin, 
  nrow = 6)
ggsave("bio_chng.tiff", plot = p, dpi = 300, width = 15, height = 30, compression = "lzw")



```

In this chunk of code, we plot the prob of occurrence based on the flow random forest.
we do bar plots to show the change in probability for high and low elv.
```{r}

head(flow)

biopred <- data.frame(
  "COMID"  = flow$COMID,
  "elv" = flow$elv,
  "ElevCat" = flow$ElevCat,
  "year" = flow$year,
  "type" = ifelse(flow$year %in% c(2095, 1993), "wet",
                  ifelse(flow$year %in% c(2010, 2100), "moderate", "dry")),
  "time" = ifelse(flow$year %in% c(2010, 1993, 2014), "baseline", "future"),
  "chub" = predict(mod_chub_rf, newdata = flow, type = "prob")[,2],
  "sucker" = predict(mod_sucker_rf, newdata = flow, type = "prob")[,2],
  "trout" = predict(mod_trout_rf, newdata = flow, type = "prob")[,2],
  "vireo" = predict(mod_vireo_rf, newdata = flow, type = "prob")[,2],
  "toad" = predict(mod_toad_rf, newdata = flow, type = "prob")[,2],
  "turtle" = predict(mod_turtle_rf, newdata = flow, type = "prob")[,2]
)

#anova to see how year type imapcts P(occurrence)
summary(aov(vireo ~ type, data=biopred))

a <- ggplot(data = biopred, mapping = aes(x = factor(elv, levels = c("low", "high")), y = chub, fill = time))+
  geom_boxplot()+ scale_fill_manual(values=c("grey90", "grey60"))+
  labs(y="P(Occurrence)", x = NULL, subtitle = "Arroyo chub")+
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5), 
        legend.title = element_blank(), 
        legend.position = c(.21, 1), 
        legend.justification = c("right", "top"),
        legend.box.background = element_rect(color="black", size=.5),
        strip.text = element_text(size = 20, face = "bold"),
        strip.background = element_rect(fill = "lightblue", colour = "black"))

b <- ggplot(data = biopred, mapping = aes(x = factor(elv, levels = c("low", "high")), y = sucker, fill = time))+
  geom_boxplot()+ scale_fill_manual(values=c("grey90", "grey60"))+
  labs(y=NULL, x = NULL, subtitle = "Santa Ana sucker")+
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5), legend.position = "none",
        strip.text = element_text(size = 20, face = "bold"),
        strip.background = element_rect(fill = "lightblue", colour = "black"))

c <- ggplot(data = biopred, mapping = aes(x = factor(elv, levels = c("low", "high")), y = trout, fill = time))+
  geom_boxplot()+ scale_fill_manual(values=c("grey90", "grey60"))+
  labs(y= NULL, x = NULL, subtitle = "Rainbow trout")+
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5), legend.position = "none",
        strip.text = element_text(size = 20, face = "bold"),
        strip.background = element_rect(fill = "lightblue", colour = "black"))

d <- ggplot(data = biopred, mapping = aes(x = factor(elv, levels = c("low", "high")), y = vireo, fill = time))+
  geom_boxplot()+ scale_fill_manual(values=c("grey90", "grey60"))+
  labs(y="P(Occurrence)", x = "Elevation", subtitle = "Least Bell's vireo")+
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5), legend.position = "none",
        strip.text = element_text(size = 20, face = "bold"),
        strip.background = element_rect(fill = "lightblue", colour = "black"))

e <- ggplot(data = biopred, mapping = aes(x = factor(elv, levels = c("low", "high")), y = toad, fill = time))+
  geom_boxplot()+ scale_fill_manual(values=c("grey90", "grey60"))+
  labs(y=NULL, x = "Elevation", subtitle = "Arroyo toad")+
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5), legend.position = "none",
        strip.text = element_text(size = 20, face = "bold"),
        strip.background = element_rect(fill = "lightblue", colour = "black"))

f <- ggplot(data = biopred, mapping = aes(x = factor(elv, levels = c("low", "high")), y = turtle, fill = time))+
  geom_boxplot()+ scale_fill_manual(values=c("grey90", "grey60"))+
  labs(y=NULL, x = "Elevation", subtitle = "Southwestern pond turtle")+
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5), legend.position = "none",
        strip.text = element_text(size = 20, face = "bold"),
        strip.background = element_rect(fill = "lightblue", colour = "black"))

margin = theme(plot.margin = unit(c(5,5,5,5), "mm"))
p<-arrangeGrob(
  a + margin, b+ margin, c+ margin, d+ margin, 
  e+ margin, f + margin, nrow = 2)
ggsave("bio_chng.tiff", plot = p, dpi = 300, width = 30, height = 15, compression = "lzw")


```

In this chunk of code, we calucate the mean change in prob of occurrence and the 5th and 95th percentiles
```{r}


#table for year type alone
bio_chg <- biopred %>%
  group_by(type) %>% 
  summarize(chub_mn = round(mean(chub[time == "future"]-chub[time == "baseline"], na.rm = T), 2),
            chub_05 = round(quantile(chub[time == "future"]-chub[time == "baseline"], na.rm = T, probs = 0.05), 2),
            chub_95 = round(quantile(chub[time == "future"]-chub[time == "baseline"], na.rm = T, probs = 0.95), 2),
            sucr_mn = round(mean(sucker[time == "future"]-sucker[time == "baseline"], na.rm = T), 2),
            sucr_05 = round(quantile(sucker[time == "future"]-sucker[time == "baseline"], na.rm = T, probs = 0.05), 2),
            sucr_95 = round(quantile(sucker[time == "future"]-sucker[time == "baseline"], na.rm = T, probs = 0.95), 2),
            trou_mn = round(mean(trout[time == "future"]-trout[time == "baseline"], na.rm = T), 2),
            trou_05 = round(quantile(trout[time == "future"]-trout[time == "baseline"], na.rm = T, probs = 0.05), 2),
            trou_95 = round(quantile(trout[time == "future"]-trout[time == "baseline"], na.rm = T, probs = 0.95), 2),
            toad_mn = round(mean(toad[time == "future"]-toad[time == "baseline"], na.rm = T), 2),
            toad_05 = round(quantile(toad[time == "future"]-toad[time == "baseline"], na.rm = T, probs = 0.05), 2),
            toad_95 = round(quantile(toad[time == "future"]-toad[time == "baseline"], na.rm = T, probs = 0.95), 2),
            vire_mn = round(mean(vireo[time == "future"]-vireo[time == "baseline"], na.rm = T), 2),
            vire_05 = round(quantile(vireo[time == "future"]-vireo[time == "baseline"], na.rm = T, probs = 0.05), 2),
            vire_95 = round(quantile(vireo[time == "future"]-vireo[time == "baseline"], na.rm = T, probs = 0.95), 2),
            turt_mn = round(mean(turtle[time == "future"]-turtle[time == "baseline"], na.rm = T), 2),
            turt_05 = round(quantile(turtle[time == "future"]-turtle[time == "baseline"], na.rm = T, probs = 0.05), 2),
            turt_95 = round(quantile(turtle[time == "future"]-turtle[time == "baseline"], na.rm = T, probs = 0.95), 2))

bio_chg <- data.frame(
  "type" = bio_chg$type,
  "chub" = paste(bio_chg$chub_mn, " (", bio_chg$chub_05, ",", bio_chg$chub_95,  ") ", sep = ""),
  "suck" = paste(bio_chg$sucr_mn, " (", bio_chg$sucr_05, ",", bio_chg$sucr_95,  ") ", sep = ""),
  "trou" = paste(bio_chg$trou_mn, " (", bio_chg$trou_05, ",", bio_chg$trou_95,  ") ", sep = ""),
  "vire" = paste(bio_chg$vire_mn, " (", bio_chg$vire_05, ",", bio_chg$vire_95,  ") ", sep = ""),
  "toad" = paste(bio_chg$toad_mn, " (", bio_chg$toad_05, ",", bio_chg$toad_95,  ") ", sep = ""),
  "turt" = paste(bio_chg$turt_mn, " (", bio_chg$turt_05, ",", bio_chg$turt_95,  ") ", sep = "")
)

bio_chg <- t(bio_chg)
write.csv(bio_chg, file = "t.csv")

```

In the final code chunk, we want a plot that shows for each spp, the direction of change for temp, and the direction of change for flow (in each wet, dry and avg year). this could be similar to the violin plot in the original paper.

```{r}

comp <- biopred %>%
  select(COMID, year, trout, chub, sucker, toad, vireo, turtle, time) %>% 
  pivot_longer(cols = 3:8, names_to = "Species", values_to = "Probability")
comp$metric <- "flow"

comp2 <- temp %>%
  select(COMID, year, trout, chub, sucker, toad, vireo, turtle, time) %>% 
  pivot_longer(cols = 3:8, names_to = "Species", values_to = "Probability") %>% 
  filter(year %in% c(1993, 2010, 2014, 2090, 2095, 2100))
comp2$metric <- "temp"
comp2$time[comp2$time == "end"] <- "future"

comp3 <- rbind(comp, comp2)
comp3$type <- ifelse(comp3$year %in% c(1993, 2095), "Wet",
                     ifelse(comp3$year %in% c(2100, 2010), "Moderate", "Dry"))

p <- ggplot(data = comp3, aes(x = factor(time), y = Probability, fill = metric))+
  geom_boxplot(width = 0.5)+scale_fill_manual(values=c("grey90", "grey40"))+
  labs(x = "Time period", y = "Probability of occurrence")+
  scale_y_continuous(breaks = seq(0, 1, by = 0.25))+
  theme(text = element_text(size=30), 
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        panel.background = element_blank(), 
        panel.border = element_rect(colour = "black", fill=NA, size=.5),
        panel.spacing.x=unit(0.5, "lines") , 
        panel.spacing.y=unit(1.5,"lines"),
        strip.text = element_text(size = 20, face = "bold"),
        strip.background = element_rect(fill = "lightblue", colour = "black"))+
  facet_wrap(vars(Species), nrow = 2)

ggsave("flow_v_temp.tiff", plot = p, dpi = 300, width = 15, height = 10, compression = "lzw")


#table for the average chang ein flow vs the average change in T
d <- comp3 %>% 
  group_by(Species, metric) %>% 
  summarise(change = round(mean(Probability[time == "future"]-Probability[time == "baseline"], na.rm = T), 2))

```