---
title: "flow"
author: "Jenny Rogers"
date: "April 8, 2020"
output: html_document
---


```{r}

library(tidyverse)
library(lubridate)
library(sf)
library(gridExtra)


```



First, we want to determine the water year types that we chose in the baseline. first group by wateryear and then sum total precip. Then calculte the percentiles of the selected baseline yeras

```{r}
load("L:/Flow ecology and climate change_ES/Marcus/flowmetrics/data/bsext.RData")

baseline_prcip <- bsext %>% 
  mutate(year = year(date), 
         month = month(date),
         water_yr = ifelse(month>=10, year+1, year)) %>% 
  group_by(water_yr) %>% 
  summarise(total = sum(dly_prp)) %>% 
  ungroup() %>% 
  mutate(rank = min_rank(total))

low <- quantile(baseline_prcip$total, probs = 0.33)
med <- quantile(baseline_prcip$total, probs = 0.66)


baseline_prcip <- baseline_prcip %>% 
  mutate(third = ifelse(total<=low, 1,
                        ifelse(total>low & total<= med, 2, 3))) 

#calucate the percentiles of the 3 years of data so we know how 3 years of data from our selected years compare
c <- baseline_prcip$total
e <- NULL
for (i in 1:(length(c)-2)){

d <- sum(c[i]+c[i+1]+c[i+2])
e <- rbind(e, d)

}
f <- rbind(NA, NA, e)
baseline_prcip$thryrsum <- f
baseline_prcip <- baseline_prcip %>% 
  mutate(rank2 = min_rank(thryrsum))

#using 3 year running sums, the percentiles are:
#1993: 0.9032258
#2010: 0.3225806
#2014: 0.03225806

```

Look at the future years to determine the annual preciptitiaon percentiles in the water years we selected

```{r}
load("L:/Flow ecology and climate change_ES/Marcus/flowmetrics/data/CanESM2ext.Rdata")
load("L:/Flow ecology and climate change_ES/Marcus/flowmetrics/data/CCSM4ext.Rdata")
load("L:/Flow ecology and climate change_ES/Marcus/flowmetrics/data/MIROC5ext.Rdata")

head(CCSM4ext)
head(CanESM2ext)
head(MIROC5ext)

#average the future daily precip projections across the three GCMs-------------------

mean_dly_prp <- (MIROC5ext$dly_prp + CanESM2ext$dly_prp + CCSM4ext$dly_prp) / 3

avg <- data.frame("year" = year(MIROC5ext$date), 
                  "month" = month(MIROC5ext$date), 
                  "day" = day(MIROC5ext$date), 
                  "COMID" = MIROC5ext$COMID, 
                  "dly_prp" = mean_dly_prp)

#Water year function
water_yr <- function(year, month){
  ifelse(month >= 10, year + 1, year)
}

avg$wateryr <- water_yr(year = avg$year, month = avg$month)

str(avg)

#total preip in each comid per year
avg <- avg %>% 
  group_by(COMID, wateryr) %>% 
  summarise(annl_prp = sum(dly_prp)) %>% 
  ungroup()

#sum the total annual precip across all locations
avg <- avg %>% 
  group_by(wateryr) %>% 
  summarise(annl_prp_sum = sum(annl_prp)) %>% 
  ungroup()

unique(avg$wateryr)

#remove the ealier decades and 1982 because only half a year
avg <- avg %>% 
  filter(wateryr >= 2083) 

plot(avg$wateryr, avg$annl_prp_sum, type = "l")

#rank the precipitation amount
avg$rank <- min_rank(avg$annl_prp_sum)

#take period of record, sort by lowest to highest, highest 3rd would be wet, middle and lower.
first_third <- quantile(avg$annl_prp_sum, probs = 0.33)
secon_third <- quantile(avg$annl_prp_sum, probs = 0.66)
avg$third <- ifelse(avg$annl_prp_sum <=first_third, 1,
                    ifelse(avg$annl_prp_sum >= first_third & avg$annl_prp_sum <= secon_third, 2, 3))

# 2095 is the wet watryear
# 2090 is the dry wateryr, 
# 2100 is the mod wateryr 


#calucate the percentiles of the 3 years of data so we know how 3 years of data from our selected years compare
c <- avg$annl_prp_sum
e <- NULL
for (i in 1:(length(c)-2)){

d <- sum(c[i]+c[i+1]+c[i+2])
e <- rbind(e, d)

}
f <- rbind(NA, NA, e)
avg$thryrsum <- f
avg <- avg %>% 
  mutate(rank2 = min_rank(thryrsum))

#using 3 year running sums, the percentiles are:
#2095: 1.0 (wettest 3 year period)
#2090: 0.0625 (driest 3 year period)
#2100: 0.375

```



In the next code chunks, we want to calutate, tabulate, and plot the change in 5 hydrologic metrics from baseline to future.

first we will read in the baseline flow metrics and the future flow metrics
```{r}
NHD <- st_read("C:/Users/JennyT/Documents/LitReview/RB4/WorkingData_3-16-18/NHDFLowline_Clip.shp")  %>%
  filter(FTYPE == "StreamRiver") %>%
  dplyr::select(COMID) %>%
  st_zm()

load("clusters.RData")

load("elv_comid.RData")


#baseline
load("bsflowmetest.RData")
baseline <- bsflowmetest %>%
  dplyr::select(var, COMID, dtsl, est) %>%
  filter(var %in% c("x3_Hydroperiod", "x3_Q99", "x3_SFR", "x3_LowDur", "x3_HighNum",
                    "all_Hydroperiod", "all_Q99", "all_SFR", "all_LowDur", "all_HighNum"),
         COMID %in% clusters$COMID) %>%
  mutate(year = year(dtsl),
         time = "baseline",
         type = ifelse(year ==1993, "wet",
                       ifelse(year == 2010, "mod", "dry"))) %>%
  spread(var, est)

#future
load("flowmetdt2.RData")
future <- flowmetdt2 %>%
  dplyr::select(var, COMID, dtsl, est) %>%
  filter(var %in% c("x3_Hydroperiod", "x3_Q99", "x3_SFR", "x3_LowDur", "x3_HighNum",
                    "all_Hydroperiod", "all_Q99", "all_SFR", "all_LowDur", "all_HighNum"),
         COMID %in% clusters$COMID) %>%
  mutate(year = year(dtsl),
         time = "future",
         type = ifelse(year ==2095, "wet",
                       ifelse(year == 2100, "mod", "dry"))) %>%
  spread(var, est)

flow <- bind_rows(baseline, future)
flow <- left_join(flow, elv, by = "COMID")

save(flow, file = "flowmet_all.RData")



```


in this next chunk of code, we will plot the distributions of COMIDs for each year type, baseline to future
```{r}

load("flowmet_all.RData")


a <- ggplot(data = flow, mapping = aes(x = factor(type, levels = c("dry", "mod", "wet")), y = x3_HighNum, fill = time))+
  geom_boxplot()+ scale_fill_manual(values=c("grey90", "grey60"))+
  labs(y="Number of Storms", x = NULL)+
  ggtitle("(A)")+ 
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5), 
        legend.title = element_blank(), legend.position = c(.225, 1),
    legend.justification = c("right", "top"),legend.box.background = element_rect(color="black", size=.5),
    legend.box.margin = margin(2, 2, 2, 2))

b <- ggplot(data = flow, mapping = aes(x = factor(type, levels = c("dry", "mod", "wet")), y = x3_LowDur, fill = time))+
  geom_boxplot()+ scale_fill_manual(values=c("grey90", "grey60"))+
  labs(y="Low Duration", x = NULL)+
  ggtitle("(C)")+ 
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5), legend.position = "none")

c <- ggplot(data = flow, mapping = aes(x = factor(type, levels = c("dry", "mod", "wet")), y = x3_Hydroperiod, fill = time))+
  geom_boxplot()+ scale_fill_manual(values=c("grey90", "grey60"))+
  labs(y="Hydroperiod", x = NULL)+
  ggtitle("(D)")+ 
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5),legend.position = "none")

d <- ggplot(data = flow, mapping = aes(x = factor(type, levels = c("dry", "mod", "wet")), y = x3_Q99, fill = time))+
  geom_boxplot()+ scale_fill_manual(values=c("grey90", "grey60"))+
  labs(y="Q99", x = NULL)+
  ggtitle("(B)")+ 
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5), legend.position = "none")

e <- ggplot(data = flow, mapping = aes(x = factor(type, levels = c("dry", "mod", "wet")), y = x3_SFR, fill = time))+
  geom_boxplot()+ scale_fill_manual(values=c("grey90", "grey60"))+
  labs(y="SFR", x = "Water Year Type")+
  ggtitle("(E)")+ 
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5), legend.position = "none")


margin = theme(plot.margin = unit(c(5,7,5,5), "mm"))
p<-arrangeGrob(
  a + margin, d+ margin, b+ margin, c + margin, e + margin,
  nrow = 5)
ggsave("flowMetChng.tiff", plot = p, dpi = 300, width = 10, height = 25, compression = "lzw")



#this is for the figure of SFR and Hydroperiod

a <- ggplot(data = flow, mapping = aes(x = factor(type, levels = c("dry", "mod", "wet")), y = x3_Hydroperiod, fill = time))+
  geom_boxplot()+ scale_fill_manual(values=c("grey90", "grey60"))+
  labs(y="Hydroperiod", x = NULL)+
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5),legend.title = element_blank(), legend.position = c(1, 1),
    legend.justification = c("right", "top"),legend.box.background = element_rect(color="black", size=.5),
    legend.box.margin = margin(1, 1, 1, 1))+
  facet_wrap(facets=vars(elv))

b <- ggplot(data = flow, mapping = aes(x = factor(type, levels = c("dry", "mod", "wet")), y = x3_SFR, fill = time))+
  geom_boxplot()+ scale_fill_manual(values=c("grey90", "grey60"))+
  labs(y="SFR", x = "Water Year Type")+
  theme(text = element_text(size=30), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5), legend.position = "none")+
  facet_wrap(facets=vars(elv))


p<-arrangeGrob(
  a + margin, b + margin,
  nrow = 2)
ggsave("flowMetChng_SFR_hydro.tiff", plot = p, dpi = 300, width = 10, height = 10, compression = "lzw")
  
  

```



Make a table of change for each year type and each elevation

```{r}
#table for year type and elevation
flow_sum <- flow %>%
  group_by(type, elv) %>% 
  summarize(HiNuChg_mn = round(mean(x3_HighNum[time == "future"]-x3_HighNum[time == "baseline"], na.rm = T), 2),
            HiNuChg_05 = round(quantile(x3_HighNum[time == "future"]-x3_HighNum[time == "baseline"], na.rm = T, probs = 0.05), 2),
            HiNuChg_95 = round(quantile(x3_HighNum[time == "future"]-x3_HighNum[time == "baseline"], na.rm = T, probs = 0.95), 2),
            LoDuChg_mn = round(mean(x3_LowDur[time == "future"]-x3_LowDur[time == "baseline"], na.rm = T), 0),
            LoDuChg_05 = round(quantile(x3_LowDur[time == "future"]-x3_LowDur[time == "baseline"], na.rm = T, probs = 0.05), 0),
            LoDuChg_95 = round(quantile(x3_LowDur[time == "future"]-x3_LowDur[time == "baseline"], na.rm = T, probs = 0.95), 0),
            Q99_mn = round(mean(x3_Q99[time == "future"]-x3_Q99[time == "baseline"], na.rm = T), 0),
            Q99_05 = round(quantile(x3_Q99[time == "future"]-x3_Q99[time == "baseline"], na.rm = T, probs = 0.05), 0),
            Q99_95 = round(quantile(x3_Q99[time == "future"]-x3_Q99[time == "baseline"], na.rm = T, probs = 0.95), 0),
            Hydrop_mn = round(mean(x3_Hydroperiod[time == "future"]-x3_Hydroperiod[time == "baseline"], na.rm = T), 2),
            Hydrop_05 = round(quantile(x3_Hydroperiod[time == "future"]-x3_Hydroperiod[time == "baseline"], na.rm = T, probs = 0.05), 2),
            Hydrop_95 = round(quantile(x3_Hydroperiod[time == "future"]-x3_Hydroperiod[time == "baseline"], na.rm = T, probs = 0.95), 2),
            SFR_mn = round(mean(x3_SFR[time == "future"]-x3_SFR[time == "baseline"], na.rm = T), 2),
            SFR_05 = round(quantile(x3_SFR[time == "future"]-x3_SFR[time == "baseline"], na.rm = T, probs = 0.05), 2),
            SFR_95 = round(quantile(x3_SFR[time == "future"]-x3_SFR[time == "baseline"], na.rm = T, probs = 0.95), 2))

flow_sum <- data.frame(
  "type" = flow_sum$type,
  "elv" = flow_sum$elv,
  "HigNum" = paste(flow_sum$HiNuChg_mn, " (", flow_sum$HiNuChg_05, ",", flow_sum$HiNuChg_95,  ") ", sep = ""),
  "LowDur" = paste(flow_sum$LoDuChg_mn, " (", flow_sum$LoDuChg_05, ",", flow_sum$LoDuChg_95,  ") ", sep = ""),
  "Q99" = paste(flow_sum$Q99_mn, " (", flow_sum$Q99_05, ",", flow_sum$Q99_95,  ") ", sep = ""),
  "Hydroperiod" = paste(flow_sum$Hydrop_mn, " (", flow_sum$Hydrop_05, ",", flow_sum$Hydrop_95,  ") ", sep = ""),
  "SFR" = paste(flow_sum$SFR_mn, " (", flow_sum$SFR_05, ",", flow_sum$SFR_95,  ") ", sep = "")
)

flow_sum <- t(flow_sum)

write.csv(flow_sum, file = "flowMetrChng_type_elv.csv")






#table for year type alone
flow_sum <- flow %>%
  group_by(type) %>% 
  summarize(HiNuChg_mn = round(mean(x3_HighNum[time == "future"]-x3_HighNum[time == "baseline"], na.rm = T), 2),
            HiNuChg_05 = round(quantile(x3_HighNum[time == "future"]-x3_HighNum[time == "baseline"], na.rm = T, probs = 0.05), 2),
            HiNuChg_95 = round(quantile(x3_HighNum[time == "future"]-x3_HighNum[time == "baseline"], na.rm = T, probs = 0.95), 2),
            LoDuChg_mn = round(mean(x3_LowDur[time == "future"]-x3_LowDur[time == "baseline"], na.rm = T), 0),
            LoDuChg_05 = round(quantile(x3_LowDur[time == "future"]-x3_LowDur[time == "baseline"], na.rm = T, probs = 0.05), 0),
            LoDuChg_95 = round(quantile(x3_LowDur[time == "future"]-x3_LowDur[time == "baseline"], na.rm = T, probs = 0.95), 0),
            Q99_mn = round(mean(x3_Q99[time == "future"]-x3_Q99[time == "baseline"], na.rm = T), 0),
            Q99_05 = round(quantile(x3_Q99[time == "future"]-x3_Q99[time == "baseline"], na.rm = T, probs = 0.05), 0),
            Q99_95 = round(quantile(x3_Q99[time == "future"]-x3_Q99[time == "baseline"], na.rm = T, probs = 0.95), 0),
            Hydrop_mn = round(mean(x3_Hydroperiod[time == "future"]-x3_Hydroperiod[time == "baseline"], na.rm = T), 2),
            Hydrop_05 = round(quantile(x3_Hydroperiod[time == "future"]-x3_Hydroperiod[time == "baseline"], na.rm = T, probs = 0.05), 2),
            Hydrop_95 = round(quantile(x3_Hydroperiod[time == "future"]-x3_Hydroperiod[time == "baseline"], na.rm = T, probs = 0.95), 2),
            SFR_mn = round(mean(x3_SFR[time == "future"]-x3_SFR[time == "baseline"], na.rm = T), 2),
            SFR_05 = round(quantile(x3_SFR[time == "future"]-x3_SFR[time == "baseline"], na.rm = T, probs = 0.05), 2),
            SFR_95 = round(quantile(x3_SFR[time == "future"]-x3_SFR[time == "baseline"], na.rm = T, probs = 0.95), 2))

flow_sum <- data.frame(
  "type" = flow_sum$type,
  "HigNum" = paste(flow_sum$HiNuChg_mn, " (", flow_sum$HiNuChg_05, ",", flow_sum$HiNuChg_95,  ") ", sep = ""),
  "LowDur" = paste(flow_sum$LoDuChg_mn, " (", flow_sum$LoDuChg_05, ",", flow_sum$LoDuChg_95,  ") ", sep = ""),
  "Q99" = paste(flow_sum$Q99_mn, " (", flow_sum$Q99_05, ",", flow_sum$Q99_95,  ") ", sep = ""),
  "Hydroperiod" = paste(flow_sum$Hydrop_mn, " (", flow_sum$Hydrop_05, ",", flow_sum$Hydrop_95,  ") ", sep = ""),
  "SFR" = paste(flow_sum$SFR_mn, " (", flow_sum$SFR_05, ",", flow_sum$SFR_95,  ") ", sep = "")
)

flow_sum <- t(flow_sum)

write.csv(flow_sum, file = "flowMetrChng_type.csv")






#table for elv type alone
flow_sum <- flow %>%
  group_by(elv) %>% 
  summarize(HiNuChg_mn = round(mean(x3_HighNum[time == "future"]-x3_HighNum[time == "baseline"], na.rm = T), 2),
            HiNuChg_05 = round(quantile(x3_HighNum[time == "future"]-x3_HighNum[time == "baseline"], na.rm = T, probs = 0.05), 2),
            HiNuChg_95 = round(quantile(x3_HighNum[time == "future"]-x3_HighNum[time == "baseline"], na.rm = T, probs = 0.95), 2),
            LoDuChg_mn = round(mean(x3_LowDur[time == "future"]-x3_LowDur[time == "baseline"], na.rm = T), 0),
            LoDuChg_05 = round(quantile(x3_LowDur[time == "future"]-x3_LowDur[time == "baseline"], na.rm = T, probs = 0.05), 0),
            LoDuChg_95 = round(quantile(x3_LowDur[time == "future"]-x3_LowDur[time == "baseline"], na.rm = T, probs = 0.95), 0),
            Q99_mn = round(mean(x3_Q99[time == "future"]-x3_Q99[time == "baseline"], na.rm = T), 0),
            Q99_05 = round(quantile(x3_Q99[time == "future"]-x3_Q99[time == "baseline"], na.rm = T, probs = 0.05), 0),
            Q99_95 = round(quantile(x3_Q99[time == "future"]-x3_Q99[time == "baseline"], na.rm = T, probs = 0.95), 0),
            Hydrop_mn = round(mean(x3_Hydroperiod[time == "future"]-x3_Hydroperiod[time == "baseline"], na.rm = T), 2),
            Hydrop_05 = round(quantile(x3_Hydroperiod[time == "future"]-x3_Hydroperiod[time == "baseline"], na.rm = T, probs = 0.05), 2),
            Hydrop_95 = round(quantile(x3_Hydroperiod[time == "future"]-x3_Hydroperiod[time == "baseline"], na.rm = T, probs = 0.95), 2),
            SFR_mn = round(mean(x3_SFR[time == "future"]-x3_SFR[time == "baseline"], na.rm = T), 2),
            SFR_05 = round(quantile(x3_SFR[time == "future"]-x3_SFR[time == "baseline"], na.rm = T, probs = 0.05), 2),
            SFR_95 = round(quantile(x3_SFR[time == "future"]-x3_SFR[time == "baseline"], na.rm = T, probs = 0.95), 2))

flow_sum <- data.frame(
  "elv" = flow_sum$elv,
  "HigNum" = paste(flow_sum$HiNuChg_mn, " (", flow_sum$HiNuChg_05, ",", flow_sum$HiNuChg_95,  ") ", sep = ""),
  "LowDur" = paste(flow_sum$LoDuChg_mn, " (", flow_sum$LoDuChg_05, ",", flow_sum$LoDuChg_95,  ") ", sep = ""),
  "Q99" = paste(flow_sum$Q99_mn, " (", flow_sum$Q99_05, ",", flow_sum$Q99_95,  ") ", sep = ""),
  "Hydroperiod" = paste(flow_sum$Hydrop_mn, " (", flow_sum$Hydrop_05, ",", flow_sum$Hydrop_95,  ") ", sep = ""),
  "SFR" = paste(flow_sum$SFR_mn, " (", flow_sum$SFR_05, ",", flow_sum$SFR_95,  ") ", sep = "")
)

flow_sum <- t(flow_sum)

write.csv(flow_sum, file = "flowMetrChng_elv.csv")

```


In this  code chunk, we want to compare the differences between the range of values in the base line and future years for each year type. for example, we want to know if the water year types are more similar in the future than in the baseline
```{r}

rng <- flow %>%
  group_by(time) %>% 
  summarise(maxHiNum = quantile(x3_HighNum, probs = 0.95, na.rm = T),
            minHiNum = quantile(x3_HighNum, probs = 0.05, na.rm = T),
            maxQ99 = quantile(x3_Q99, probs = 0.95, na.rm = T),
            minQ99 = quantile(x3_Q99, probs = 0.05, na.rm = T),
            maxLoDur = quantile(x3_LowDur, probs = 0.95, na.rm = T),
            minLoDur = quantile(x3_LowDur, probs = 0.05, na.rm = T),
            maxHydro = quantile(x3_Hydroperiod, probs = 0.95, na.rm = T),
            minHydro = quantile(x3_Hydroperiod, probs = 0.05, na.rm = T),
            maxSFR = quantile(x3_SFR, probs = 0.95, na.rm = T),
            minSFR = quantile(x3_SFR, probs = 0.05, na.rm = T)) %>% 
  ungroup()
rng <- rng %>% 
  mutate(HiNumRng = maxHiNum - minHiNum,
         maxQ99Rng = maxQ99 - minQ99,
         LoDurRng = maxLoDur - minLoDur,
         HydroRng = maxHydro - minHydro,
         SFRRng = abs(maxSFR - minSFR))

prctRngChng <- rng[2,12:16]/rng[1,12:16]

```

In this code chunk, we copare the similarly of year types

```{r}
similarly <- flow %>%
  group_by(time, type) %>% 
  summarize(muHN = round(mean(x3_HighNum, na.rm = T), 2),
            sdNH  = round(sd(x3_HighNum, na.rm = T), 2),
            muLD = round(mean(x3_LowDur, na.rm = T), 0),
            sdLD  = round(sd(x3_LowDur, na.rm = T), 0),
            mu99 = round(mean(x3_Q99, na.rm = T), 0),
            sd99 = round(sd(x3_Q99, na.rm = T), 0),
            muSFR = round(mean(x3_SFR, na.rm = T), 2),
            sdSFR = round(sd(x3_SFR, na.rm = T), 2),
            muhyd = round(mean(x3_Hydroperiod, na.rm = T), 2),
            sdhyd = round(sd(x3_Hydroperiod, na.rm = T), 2))

similarly <- data.frame(
  "time" = similarly$time,
  "type" = similarly$type,
  "highNum" = paste(similarly$muHN, "\u00B1", similarly$sdNH),
  "Q99" = paste(similarly$mu99, "\u00B1", similarly$sd99),
  "LoDur" = paste(similarly$muLD, "\u00B1", similarly$sdLD),
  "Hydroperiod" = paste(similarly$muhyd, "\u00B1", similarly$sdhyd),
  "SFR" = paste(similarly$muSFR, "\u00B1", similarly$sdSFR)
)

similarly <- t(similarly)
write.csv(similarly, file = "YearTypeSimilarly_nowVsFuture.csv")

```


In this code chunk, we are going to subtract the baseline values from the future values for each variable and run a paired t-test to determine if the change is different than zero.  need to change the variable in the initial select and spread, and then nothing else needs to change.

```{r}


hyd <- flow %>% 
  select(COMID, time, type, x3_SFR, elv) %>% 
  spread(time, x3_SFR) 

d <-  hyd %>% 
  filter(elv == "high" & type == "dry")
t.test(d$future, d$baseline, paired = TRUE)

d <-  hyd %>% 
  filter(elv == "high" & type == "mod")
t.test(d$future, d$baseline, paired = TRUE)

d <-  hyd %>% 
  filter(elv == "high" & type == "wet")
t.test(d$future, d$baseline, paired = TRUE)

d <-  hyd %>% 
  filter(elv == "low" & type == "dry")
t.test(d$future, d$baseline, paired = TRUE)

d <-  hyd %>% 
  filter(elv == "low" & type == "mod")
t.test(d$future, d$baseline, paired = TRUE)

d <-  hyd %>% 
  filter(elv == "low" & type == "wet")
t.test(d$future, d$baseline, paired = TRUE)


```